---
title: "Apex_Water_Chem_Nov12-29"
author: "Shelly Trigg"
date: "11/28/2018"
output: rmarkdown::github_document
---


```{r}
#load libraries
library("XML")
library("plyr")
library("tidyr")
library("ggplot2")
```

```{r, eval = FALSE}
#read in the date plus x days of Apex data
xmlfile <- xmlParse("http://192.168.1.100:80/cgi-bin/datalog.xml?sdate=181107&days=21")

#convert xml to dataframe
Apex.Data <- ldply(xmlToList(xmlfile), data.frame)

#output the Apex data to a csv file; this is to make a copy of the raw data
write.csv(Apex.Data, "~/Documents/GitHub/P_generosa/Water_Chemistry/data/Apex/Apex_data_20181107-20181128.csv", quote =FALSE, row.names = FALSE)
```

```{r}
#read in the csv file
Apex.Data <- read.csv("~/Documents/GitHub/P_generosa/Water_Chemistry/data/Apex/Apex_data_20181107-20181128.csv", stringsAsFactors = FALSE)

#remove uninformative rows and columns
Apex.Data <- Apex.Data[-c(1:3),-c(1:2,grep("type", colnames(Apex.Data)))]

#make a data frame of just probe names
Apex.Data.probe <- Apex.Data[,c(1,grep("name", colnames(Apex.Data)))]
#make a data frame of just probe values
Apex.Data.val <- Apex.Data[,c(1,grep("value", colnames(Apex.Data)))]
#reshabe data frames into long format
Apex.Data.probe <- gather(Apex.Data.probe, probe_name, value ,-1)
Apex.Data.val <- gather(Apex.Data.val, value2, value ,-1)
#recombine probe name and value long data frames
Apex.Data <- cbind(Apex.Data.probe[,-2], Apex.Data.val[,3])
#extract pH, salt, oxygen, and temperature data
Apex.Data <- Apex.Data[grep("pH|Tmp|Salt|ORP", Apex.Data$value),]
#convert date/time from character to POSIX so R will read it as a date
Apex.Data$date <- as.POSIXct(strptime(Apex.Data$date, "%m/%d/%Y %H:%M:%S"))
#rename columns
colnames(Apex.Data) <- c("date", "probe", "value")
#Probe 'Tmpx6' and 'pHx6' were renamed to 'Tmp-2' and 'pH-2' so rename them in the data to reflect that
Apex.Data$probe <- gsub("x6", "_2", Apex.Data$probe)
#replace the hyphen in 'Tmp-2' and 'pH-2' with an underscore
Apex.Data$probe <- gsub("-","_", Apex.Data$probe)
#Probe 'Tmp' was renamed to 'Tmp_1' so rename it in the data to reflect that
Apex.Data$probe <- gsub("^Tmp$", "Tmp_1", Apex.Data$probe)
#Probe 'pH' was renamed to 'pH_1' so rename it in the data to reflect that
Apex.Data$probe <- gsub("^pH$", "pH_1", Apex.Data$probe)
```

Plotting daily pH variability in ambient treatments
Feeding typically happens sometime between 8-10am and 4-6pm. The tanks are cleaned (emptied, sprayed, and refilled) on Fridays (Nov. 2 and 9) and possibly some other days. Apex records data every 10 minutes.
```{r, echo= FALSE}
#pull out only data for Tanks 3 and 4 (ambient treatment)
#exlude data before Nov 7 for Tank 3 since this probe was actually in tank2 at that time
#exclude Oct 26 data because that was day one data and wasn't stable
amb <- Apex.Data[which(substr(Apex.Data$probe,0,4)=="pH_3" & Apex.Data$date > strptime("11/07/2018 00:00:00", format = "%m/%d/%Y %H:%M:%S") | substr(Apex.Data$probe,0,4)=="pH_4" & Apex.Data$date > strptime("10/27/2018 00:00:00", format = "%m/%d/%Y %H:%M:%S")),]
#exclude calibrations
amb <- amb[which(amb$value < 9 & amb$value > 6),]

#make a column with only the time
amb$time <- strftime(amb$date, format = "%H:%M:%S")
#make a column for day and format it so it has the abbv. month and day
amb$day <- format(amb$date, "%b %d")
```

### Daily pH variability 
```{r, fig.width = 11, fig.height = 5, echo = FALSE}

ggplot(amb, aes(x= date, y= value, color = probe)) + geom_point() +scale_x_datetime("date", date_breaks = "day", date_labels = "%b %d", expand = c(0,0), minor_breaks = NULL)  + xlab("date") + ylab("pH") + theme(axis.text.x=element_text(angle = 90, hjust = 0)) + ggtitle("pH over time for ambient tanks 3 and 4")
```
I excluded feeding times to get a better sense of daily pH variability. 
```{r, fig.width = 13, fig.height = 5, echo = FALSE}
### Daily pH variability excluding feeding times
#exclude feeding times
amb_nofeed <- rbind(amb[which(amb$time < "08:00:00"),], amb[which(amb$time > "11:00:00" & amb$time < "15:00:00"),], amb[which(amb$time > "18:00:00"),])

ggplot(amb_nofeed, aes(x= date, y= value, color = probe)) + geom_point() +scale_x_datetime("date", date_breaks = "day", date_labels = "%b %d", expand = c(0,0), minor_breaks = NULL)  + xlab("date") + ylab("pH") + theme(axis.text.x=element_text(size = 5, angle = 90, hjust = 0)) + ggtitle("pH over time EXCLUDING FEEDING TIMES")
```

```{r, fig.width = 13, fig.height = 15, echo = FALSE}
ggplot(amb, aes(x= time, y= value, color = probe)) + geom_point() + xlab("time") + ylab("pH")  + 
  facet_wrap(~day, ncol = 2, scale = "free") + theme(axis.text.x=element_text(size = 7, angle = 90, hjust = 0)) + ggtitle("daily pH over time for ambient tanks 3 and 4")
```

```{r, fig.width = 13, fig.height = 15, echo = FALSE}
ggplot(amb_nofeed, aes(x= time, y= value, color = probe)) + geom_point() + xlab("day") + ylab("pH")  + 
  facet_wrap(~day, nrow = 6, scale = "free") + theme(axis.text.x=element_text(size = 5, angle = 90, hjust = 0)) + ggtitle("daily pH over time for ambient tanks 3 and 4 EXCLUDING FEEDING TIMES")
```


```{r, fig.width = 13, fig.height = 5, echo = FALSE}
ggplot(amb[which(amb$date > strptime("11/07/2018 00:00:00", format = "%m/%d/%Y %H:%M:%S")),], aes(time, value)) + geom_boxplot(aes(fill = value, color = probe)) + xlab("day") + ylab("pH") + theme(axis.text.x=element_text(angle = 90, hjust = 0)) + ggtitle("boxplots of daily pH variability from Nov 7-12")
```
```{r, fig.width = 13, fig.height = 5, echo = FALSE}
ggplot(amb_nofeed[which(amb_nofeed$date > strptime("11/07/2018 00:00:00", format = "%m/%d/%Y %H:%M:%S")),], aes(time, value)) + geom_boxplot(aes(fill = value, color = probe)) + xlab("day") + ylab("pH") + theme(axis.text.x=element_text(angle = 90, hjust = 0)) + ggtitle("boxplots of daily pH variability from Nov 7-12 EXCLUDING FEEDING TIMES")